---
title: "Extra Dataset Report"
author: "LAM"
date: "4/14/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, out.width='50%', fig.align = "center")

library(dplyr)
library(magrittr)
library(factoextra)

pca_fun <- function(dat){
pca <- prcomp(select_if(dat, is.numeric),
              center = TRUE,
              scale = TRUE)
}

scree_plot <- function(pca){
factoextra::fviz_eig(pca,
  geom = c("bar"),
  addlabels = T
) +
  geom_line(aes(x = , y = head(get_eig(pca),10)$cumulative.variance.percent),)
}

# Create loading plot
loading_plot <- function(pca){
  
  factoextra::fviz(
  pca,
  element = "var",
  title = "Loading plot",
  geom = c("point", "text")
)

}

# Create score plot
score_plot <- function(pca) {
  
  factoextra::fviz(pca,
                 element = "ind",
                 # Using "." makes the plotting faster
                 pointshape = ".",
                 geom = "point",
                 title = "Score plot",
                 addEllipses = T,
                 ellipse.level = 0.95) +
  scale_x_continuous(limits = c(-10, 16)) +
  labs(caption = "Conf. Ellipse - 95%")
}
```

## Data report

### MHEALTH

```{r}
dat <- read.csv(here::here("data/MHEALTH_cleaned.csv")) %>%
  select_if(is.numeric) %>%
  filter(complete.cases(.))
```

The MHEALTH (Mobile Health) dataset is devised to benchmark techniques dealing with human behavior analysis based on multimodal body sensing. It contains `r ncol(dat)` numeric variables, and has `r nrow(dat)` complete observations.

```{r, out.width='70%'}
pca <- dat %>% 
  pca_fun()

scree_plot(pca)
```

```{r, fig.align='default', fig.show = 'hold'}
score_plot(pca)
loading_plot(pca)
```

\newpage

## Dow Jones dataset

```{r}
dat <- read.csv(here::here("data_raw/dow_jones_index.data")) %>%
  select_if(is.numeric) %>%
  filter(complete.cases(.))
```

This dataset contains weekly data for the Dow Jones Industrial Index. It contains `r ncol(dat)` numeric variables, and has `r nrow(dat)` complete observations.

```{r, out.width='70%'}
pca <-  dat %>%
  pca_fun()

scree_plot(pca)
```

```{r, fig.align='default', fig.show = 'hold'}
score_plot(pca)
loading_plot(pca)
```

\newpage

## City power consumption

```{r}
dat <- read.csv(here::here("data_raw/Tetuan_City_power_consumption.csv")) %>%
  select_if(is.numeric) %>%
  filter(complete.cases(.))
```

This dataset is related to power consumption of three different distribution networks of Tetouan city which is located in north Morocco. It contains `r ncol(dat)` numeric variables, and has `r nrow(dat)` complete observations.

```{r, out.width='70%'}
pca_power <- dat %>%
  pca_fun()

scree_plot(pca_power)
```

```{r, fig.align='default', fig.show = 'hold'}
score_plot(pca_power)
loading_plot(pca_power)
```

\newpage

## Metro Interstate Traffic Volume

```{r}
dat <- read.csv(here::here("data_raw/Metro_Interstate_Traffic_Volume.csv.gz")) %>%
  select_if(is.numeric) %>%
  filter(complete.cases(.))
```

This dataset describes hourly Minneapolis-St Paul, MN traffic volume for westbound I-94. *Note:* this dataset contains information on weather events and so could give a natural subset of observations (e.g. only want to examine traffic volume when it is raining). This dataset contains `r ncol(dat)` numeric variables, and has `r nrow(dat)` complete observations.

```{r, out.width='70%'}
pca_power <- dat %>%
  pca_fun()

scree_plot(pca_power)
```

```{r, fig.align='default', fig.show = 'hold'}
score_plot(pca_power)
loading_plot(pca_power)
```

\newpage


